shrimp_goby_boatnoise

### 2025-02-18 ms_boatnoise_shrimp_goby.Rmd
adding code

remotes::install_github("stan-dev/loo")
# Setting colours for vidsalisations
colour1 <- "#97A275"
colour2 <- "#F2B23A"
colour3 <- "#F2711B"
colour4 <- "#8CBEB2"
colour5 <- "#E7C27D"
colour6 <- "#F2A775"
colour11 <- "#666E4F"
colour12 <- "#A38346"
colour13 <- "#9C4811"
```

@@

#reading in the coded raw BORIS output
BORIS_output <- read.csv("./2-data/BORIS_output_update.csv")

@@

#renaming some of the variables and extracting individual variables from columns that have multiple variables contained in one string
decoded_dataframe <- decoded_dataframe |> 
  dplyr::rename(stage = video_section,
                duration = total_duration_s,
                behaviour = behavior) |>
  dplyr::mutate(date = str_sub(date_file, , 6),
                species = str_sub(subject, 5,),
                duration = as.numeric(duration),
                boat_id = str_sub(date_file, -1, -1),
                video_num = as.numeric(str_sub(video_id, 2)),
                burrow = ceiling(video_num/3),
                individual = str_sub(subject, 1, 1),
                helper = str_sub(date_file, -6,-6),
                stroke = ifelse(helper == "t", 'control', 
                                ifelse((boat_id == "A")|(boat_id == "B")|(boat_id == "C")|(boat_id == "D"), 
                                       '2', '4')))

@@

final_df <- inner_join(goby_df, shrimp_df, by = "burrow_stage") |> 
  mutate(shrimp_contact = ifelse(is.na(shrimp_contact), ifelse(goby_out == 1, 0, 1 - goby_out), shrimp_contact))

@@

#2. Statistical analysis using zero-one inflated beta regression
##2.1 Building a global model
```{r}
#building a global model for the proportion of time gobies spent out of the burrow
zoib_formula <- bf(
  # mu (mean) part
  goby_out ~ stroke*stage + goby_number + shrimp_number + shrimp_species + (1 | site/burrow),
  # phi (precision) part
  phi ~ stroke*stage + goby_number + shrimp_number + shrimp_species + (1 | site/burrow),
  # alpha (zero-one-inflation) part
  zoi ~ stroke*stage + goby_number + shrimp_number + shrimp_species + (1 | site/burrow),
  # The one-inflated part, conditional on the 0s
  coi ~ stroke*stage + goby_number + shrimp_number + shrimp_species + (1 | site/burrow),

@@

goby_out_model_estimates

write.csv(goby_out_model_estimates, "./2-data/goby_out_model_estimates.csv")

@@

goby_out_model_estimates_two <- goby_out_model_estimates |> 
  dplyr::filter(stroke == "2")|> 
  rename(.value = emmean)
```
####2.3.3.2 Plotting model estimates and raw data
```{r}
goby_out_model_draws <- goby_out_model |> 
  emmeans(~ stroke*stage,
          epred = TRUE,
          rg.limit = 14000,
          re_formula = NA) |> 
  gather_emmeans_draws()

@@

goby_out_fig_together
ggsave("./3-figs/goby_out_fig_together.png", plot = goby_out_fig_together, units = "px", width = 2200, height = 1400, dpi = 300)
```

@@

goby_out_shrimp_species

@@

goby_out_shrimp_species_estimates

@@

###2.4.2 Posterior marginal effects
####2.4.2.1 Pairwise contrasts of stroke and stage
```{r}
ranef(shrimp_out_model)

shrimp_out_contrasts <- shrimp_out_model |>
  emmeans(~ stroke*stage, 
          epred = TRUE,
          re_formla = NULL,
          rg.limit =14000) |> 
  contrast(method = "revpairwise")

@@

write.csv(shrimp_out_model_estimates, "./2-data/shrimp_out_model_estimates.csv")

@@

shrimp_out_fig_together
ggsave("./3-figs/shrimp_out_fig_together.png", plot = shrimp_out_fig_together, units = "px", width = 2200, height = 1400, dpi = 300)
```

@@

shrimp_out_shrimp_number_contrasts

@@

@@

write.csv(shrimp_contact_model_estimates, "./2-data/shrimp_contact_model_estimates.csv")

@@

>>>>

#remotes::install_github("stan-dev/loo")
# Setting colours for vidsalisations
colour1 <- "#97A275"
colour2 <- "#F2B23A"
colour3 <- "#F2711B"
colour4 <- "#8CBEB2"
colour5 <- "#E7C27D"
colour6 <- "#F2A775"
colour11 <- "#666E4F"
colour12 <- "#A38346"
colour13 <- "#9C4811"
```

@@

#reading in the coded raw BORIS output
BORIS_output <- read.csv("./2-data/BORIS_output.csv")

@@

#renaming some of the variables and extracting individual variables from columns that have multiple variables contained in one string
decoded_dataframe <- decoded_dataframe |> 
  dplyr::rename(stage = video_section,
                duration = total_duration_s,
                behaviour = behavior) |>
  dplyr::mutate(date = str_sub(date_file, , 6),
                species = str_sub(subject, 5,),
                duration = as.numeric(duration),
                boat_id = str_sub(date_file, -1, -1),
                video_num = as.numeric(str_sub(video_id, 2)),
                burrow = ceiling(video_num/3),
                individual = str_sub(subject, 1, 1),
                stroke = factor(stroke),
                boat_name = factor(boat_name)
                )

@@

final_df <- inner_join(goby_df, shrimp_df, by = "burrow_stage") |> 
  mutate(shrimp_contact = ifelse(is.na(shrimp_contact), ifelse(goby_out == 1, 0, 1 - goby_out), shrimp_contact)) |> 
  filter(site != "SITE O")

@@

#2. Statistical analysis using zero-one inflated beta regression
##2.1 Building a global model
```{r}
#building a global model for the proportion of time gobies spent out of the burrow
zoib_formula <- bf(
  # mu (mean) part
  goby_out ~ stroke * stage + goby_number + shrimp_number + shrimp_species + (1 | site/burrow),
  # phi (precision) part
  phi ~ stroke * stage + goby_number + shrimp_number + shrimp_species + (1 | site/burrow),
  # alpha (zero-one-inflation) part
  zoi ~ stroke * stage + goby_number + shrimp_number + shrimp_species + (1 | site/burrow),
  # The one-inflated part, conditional on the 0s
  coi ~ stroke * stage + goby_number + shrimp_number + shrimp_species + (1 | site/burrow),

@@

@@

goby_out_model_estimates_two <- goby_out_model_estimates |> 
  dplyr::filter(stroke == "2")|> 
  rename(.value = emmean)
```

####2.3.3.2 Plotting model estimates and raw data
```{r}
goby_out_model_draws <- goby_out_model |> 
  emmeans(~ stroke*stage,
          epred = TRUE,
          rg.limit = 14000,
          re_formula = NA) |> 
  gather_emmeans_draws()

@@

goby_out_fig_together
```

@@

goby_out_shrimp_species <- as.data.frame(goby_out_shrimp_species)

@@

goby_out_shrimp_species_estimates <- as.data.frame(goby_out_shrimp_species_estimates)

@@

###2.4.2 Posterior marginal effects
####2.4.2.1 Pairwise contrasts of stroke and stage
```{r}
shrimp_out_contrasts <- shrimp_out_model |>
  emmeans(~ stroke*stage, 
          epred = TRUE,
          re_formla = NULL,
          rg.limit =14000) |> 
  contrast(method = "revpairwise")

@@

@@

shrimp_out_fig_together
```

@@

shrimp_out_shrimp_number_contrasts <- as.data.frame(shrimp_out_shrimp_number_contrasts)

@@

overall_mean <- shrimp_contact_model |>
  emmeans(~ 1,             # 1 means "no factors," i.e. the grand mean
          epred = TRUE,    # estimate on the response scale
          re_formla = NULL,
          rg.limit = 14000)

# View the summary with posterior mean and credible intervals
overall_mean

@@

@@

```{r}
table_S1 <- final_df %>%
  select(c("burrow", "stroke", "site")) |> 
  unique() |> 
  group_by(site, stroke) %>%
  summarise(n = n(), .groups = "drop") %>%
  pivot_wider(names_from = stroke, values_from = n, values_fill = 0) %>%
  mutate(Total = `control` + `4` + `2`) %>%
  arrange(site) %>%
  rename(`4-stroke` = `4`, 
         `2-stroke` = `2`, 
         Control = control)
```
